---
title: "A Place For My Books"

output: rmdformats::readthedown
---

```{r setup, echo = FALSE, message = FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
library(tidyverse)
library(DT)
library(googlesheets4)
library(kableExtra)
library(lubridate)
library(janitor)

allbooks <- read_sheet(ss = '1_qKs7DbC-8uJoCV5WfgvDMLSOFKgKH7VRDWSsYch540') %>% as_tibble() %>%  mutate_at(c("Published","Started","Finished"),ymd) 

bestbooks <- select(filter(allbooks,!is.na(Category)),c(Title,Author,Published,Started,Finished,Pages,FicNonfic,Genre,Series,Category,BestCat,LinkedTitle,Image,Description)) 

booklist<-select(filter(allbooks, !is.na(Started)),c(Title,Author,Published,Started,Finished,Pages,FicNonfic,Genre,Country))

current_list <- select(filter(allbooks,!is.na(Last5)),c(Title,Author,Published,Started,Finished,Pages,FicNonfic,Genre,Series,Category,BestCat,LinkedTitle,Image,Description,Last5))

unread <- allbooks[str_detect(allbooks$Order,"U"),] %>% select(c(Title,Author,Published,Pages,FicNonfic,Genre))

image_table <- function(text_tbl,image_tbl) {
  kbl(text_tbl) %>%
  kable_paper(full_width = F) %>%
  column_spec(1, image = spec_image(image_tbl$Image,124,200), width = "10%") %>%
  column_spec(2, width = "15%") %>%
  column_spec(3, width = "15%") %>%
  column_spec(4, width = "60%") %>%
  kable_classic(full_width = F, html_font = "Cambria")
}

author_table <- function(author) {
  author_data <- bestbooks %>% filter(str_detect(Author,author)) %>% filter(str_detect(BestCat,"Authors")) %>% arrange(Published)
text_tbl <- data.frame(
  Cover ='',
  Book = author_data$LinkedTitle,
  Author = author_data$Author,
  Description = author_data$Description)
image_table(text_tbl,author_data)
}

series_table <- function(series_name) {
  series_data <- bestbooks %>% filter(str_detect(Series,series_name)) %>% filter(str_detect(BestCat,"Series")) %>% arrange(Published)
  text_tbl <- data.frame(
  Covers ='',
  Book = series_data$LinkedTitle,
  Author = series_data$Author,
  Description = series_data$Description)
image_table(text_tbl,series_data)
}

year_table <- function(yearbest) {
  year_data <- bestbooks %>% filter(str_detect(BestCat,yearbest)) %>% arrange (Finished)
  text_tbl <- data.frame(
  Covers ='',
  Book = year_data$LinkedTitle,
  Author = year_data$Author,
  Description = year_data$Description)
image_table(text_tbl,year_data)
}

top_25_authors <- function() {
  top_data <- bestbooks %>% filter(str_detect(Category,"Authors")) %>% arrange(Author,Published)
  text_tbl <- data.frame(
    Book = top_data$Title,
    Author = top_data$Author)
  kbl(text_tbl) %>%
  kable_paper(full_width = F) %>%
  column_spec(1, width = "50%") %>%
  column_spec(2, width = "50%") %>%
  kable_classic(full_width = F, html_font = "Cambria")
}

top_25_series <- function(seriestype) {
  top_data <- bestbooks %>% filter(str_detect(Category,seriestype)) %>% arrange(Series,Published)
  text_tbl <- data.frame(
    Book = top_data$Title,
    Author = top_data$Author,
    Series = top_data$Series)
  kbl(text_tbl) %>%
  kable_paper(full_width = F) %>%
  column_spec(1, width = "35%") %>%
  column_spec(2, width = "35%") %>%
  column_spec(3, width = "30%") %>%  
  kable_classic(full_width = F, html_font = "Cambria")
}
top_25_other <- function(cat_choice) {
  top_data <- bestbooks %>% filter(str_detect(Category,cat_choice)) %>% filter(!str_detect(Category,"Series")) %>% arrange(Title)
  text_tbl <- data.frame(
    Book = top_data$Title,
    Author = top_data$Author
    )
  kbl(text_tbl) %>%
  kable_paper(full_width = F) %>%
  column_spec(1, width = "50%") %>%
  column_spec(2, width = "50%") %>%
  kable_classic(full_width = F, html_font = "Cambria")
}

latest_books <- function(timing) {
  book_data <- current_list %>% filter(str_detect(Last5,timing)) %>%  arrange(Started)
text_tbl <- data.frame(
  Cover ='',
  Book = book_data$LinkedTitle,
  Author = book_data$Author,
  Description = book_data$Description)
image_table(text_tbl,book_data)
}

stat_summary <- function(year) {
  totals_year = booklist %>% filter(year(Finished) == year) %>% group_by(FicNonfic) %>% dplyr::summarise(PageCount = sum(Pages),BookCount = n_distinct(Title)) %>% adorn_totals("row", fill = "Total") %>% mutate (PageCount = formatC(round(PageCount),format = "f", big.mark = ",", drop0trailing = TRUE))
datatable(totals_year)
}

```
# Welcome
As an avid reader and an avid data nerd, I decided to create this site to track some of my reading and favourite books, both for my own interest and for anyone looking for good book recommendations. Links where applicable are to independent bookstores. Support libraries and independent bookstores!

## Current Reading {.tabset .tabset-pills}
### Current Books Reading
These are the books I'm reading right now. I often have an e-book from the library, a paper book from either my shelf or the library, and an active read-aloud book.
```{r}
latest_books("Current")
```
### Last 5 Books Read
These are the last 5 books that I've finished.
```{r}
latest_books("Last")
```
# Best Books of 2020 to 2022
The books in these sections are all ones that I've read since the start of 2020 (not necessarily books published in those years). I have put the authors whose works of fiction I've read and enjoyed at least 5 of (separate from single series) in their own categories here, because they would take up a lot of the different years' top books. I've also enjoyed reading a number of science fiction and fantasy series. Rather than choose individual ones for best-of-year lists, I'm just listing my favourite series in whole here on their own.  After that are my favourite primarily standalone works in fiction or non-fiction that I read in each of those years. They are ordered by when I read them during the year.


## Favourite Authors {.tabset .tabset-pills}
### Richard Wagamese
```{r}
author_table("Wagamese")
```

### Fredrik Backman
```{r}
author_table("Backman")
```
### Becky Chambers
```{r}
author_table("Chambers")
```
### Thomas King
```{r}
author_table("Thomas King")
```

## Favourite Sci-Fi / Fantasy Series {.tabset .tabset-pills}
### The Expanse
```{r}
series_table("Expanse")
```
### The Murderbot Diaries
```{r}
series_table("Murderbot")
```
### The Broken Earth Trilogy
```{r}
series_table("Broken Earth")
```
### The Oxford Time Travel Series
```{r}
series_table("Oxford")
```
### Remembrance of Earth's Past
```{r}
series_table("Remembrance")
```
### Hyperion Cantos
```{r}
series_table("Hyperion")
```
### The Imperial Radch
```{r}
series_table("Radch")
```
### Teixcalaan
```{r}
series_table("Teixcalaan")
```
### The Machineries of Empire
```{r}
series_table("Machineries")
```

## 2022 {.tabset .tabset-pills}
### Top 10 Fiction
```{r}
year_table("2022F")
```
### Top 10 Non-Fiction
```{r}
year_table("2022N")
```
### Stats
```{r}
stat_summary(2022)
```
## 2021 {.tabset .tabset-pills}
### Top 10 Fiction
```{r}
year_table("2021F")
```
### Top 10 Non-Fiction
```{r}
year_table("2021N")
```
### Stats
```{r}
stat_summary(2021)
```
## 2020 {.tabset .tabset-pills}
### Top 10 Fiction
```{r}
year_table("2020F")
```
### Top 10 Non-Fiction
```{r}
year_table("2020N")
```
### Stats
```{r}
stat_summary(2020)
```
# Top 25's

Here I went further back, as my book tracking includes back to 2015, and I also pulled in some books I've read over years before that, and tried to make an active list of top 25 books in each category, regardless of when I read them. There is lots of overlap with the 3 years above due to the number of books I read.  A note on earth-based and non-earth based is simply that some of the sci-fi series were set on Earth or specified Earth-origin characters, and some were not. This made a reasonable way to split them in half rather than having a list of 50 books.

## Fiction {.tabset .tabset-pills}
### Favourite Authors
```{r}
top_25_authors()
```
### Earth-based Spec Fic 
```{r}
top_25_series("SeriesEarth")
```
### Non-Earth Spec Fic Series
```{r}
top_25_series("SeriesOther")
```
### Speculative Fiction
```{r}
top_25_other("Speculative")
```
### Contemporary Fiction
```{r}
top_25_other("Contemporary")
```
### Historical Fiction
```{r}
top_25_other("Historical")
```

## Non-Fiction {.tabset .tabset-pills}
### History / Memoir / Narrative
```{r}
top_25_other("History")
```
### Social Justice
```{r}
top_25_other("Justice")
```
### Science / Nature
```{r}
top_25_other("Science")
```
### Other Non-Fiction
```{r}
top_25_other("Other")
```
# Full Lists

Since I've been actively tracking my reading (2015 and on), below are the fiction and non-fiction I've read in that time.  The lists are sortable and searchable. Following that are books I have on my shelf. They're separated into unread and read, and can be searched and sorted.

## Reading List {.tabset .tabset-pills}
### Fiction read since 2015
```{r}
booklist %>% filter(!str_detect(FicNonfic,"Non")) %>% datatable()
```

### Non-Fiction read since 2015
```{r}
booklist %>% filter(str_detect(FicNonfic,"Non")) %>% datatable()
```

## Personal Bookshelf {.tabset .tabset-pills}
### Unread
```{r}
select(filter(allbooks,(is.na(Started) | Order == "U") & Owned == 1),c(Title,Author,Pages,FicNonfic,Genre,Published)) %>% arrange(Title) %>% datatable()
```

### Read
```{r}
select(filter(allbooks,(!is.na(Finished) | Order == "R") & Owned == 1),c(Title,Author,Pages,FicNonfic,Genre,Published)) %>% arrange(Title) %>% datatable()
```

# Looking for Book Recommendations?
These are some of the sites I look at for book reviews and lists of books.  
* [Literary Hub](https://lithub.com/) - Recommendations and reviews over all categories  
* [Book Riot](https://bookriot.com/) - Recommendations and reviews over all categories  
* [Tor.com](https://www.tor.com/) - Science Fiction and Fantasy  
* [Book Marks](https://bookmarks.reviews/) - Aggregated reviews of books  
* [NPR Books](https://apps.npr.org/best-books/#view=covers&year=2022) - NPR's favourite books of the year (2022-2013), sortable by many categories  
* [Literature Map](https://www.literature-map.com/) - Put in an author you like, and find a bunch of new ones you might enjoy  
* [Five Books](https://fivebooks.com/) - Top 5's in a bunch of categories  
* [Electric Literature](https://electricliterature.com/) - Reading lists, articles, essays, and more  